Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance and associated polyandry

### 2017-03-18 ms_pre_post.Rmd
Edits up to Discussion

@@

<!--- Last time we were through this, it was suggested that I scrap the former term ''inbreeding strategy'' because we don't deal at all with inbreeding preference. I have rewritten these sections in attempt to make the writing clearer. TODO: Additionally, I have tweaked the figure legends appropriately. --->

However, when both pre-copulatory and post-copulatory inbreeding avoidance could evolve, their relative evolutionary dynamics depended on their relative costs. When pre-copulatory but not post-copulatory inbreeding avoidance was costly ($c_{M}=0.02$ and $c_{F}=0$), pre-copulatory inbreeding avoidance initially evolved (i.e., negative $M_{a}$) but then evolved back towards random mating (i.e., $M_{a}=0$) following increasing evolution of post-copulatory inbreeding avoidance (Figure 3B). Similarly, when post-copulatory but not pre-copulatory inbreeding avoidance was costly ($c_{F}=0.02$ and $c_{M}=0$), post-copulatory inbreeding avoidance intially evolved ($F_{a}=0$) before returning to random fertilisation after ca $20000$ generations (Figure 3D).

<!--- Something about this 'evolved' and 'evolved back' language just sounds very amateurish -- can't we reframe this in terms of selection? --->

Evolution of post-copulatory inbreeding avoidance facilitated evolution of polyandry (Figure 3B,C). In the absence of post-copulatory inbreeding avoidance, polyandry did not evolve (Figure 3A,D; polyandry was cost free in all these simulations).

@@

<!--- Mean pre-cop value is -17.62444  in Figure 1B --->
<!--- Mean pre-cop value is -15.44786  in Figure 4A --->

<!--- I know I've mentioned this too many times, but I'm really not a fan of this kind of quantitative statistic because it doesn't really mean anything in the context of a model like this. At the very least though, I think it's better to put the actual numbers up because ''X% less negative'' just seems very confusing to me. --->

<!--- I added this last sentence for summary, but stopped it a bit earlier than the suggestion because I'm not sure where you were going with it. If we know how selection is changed, isn't this the point? Evolution more broadly encompasses mutation and drift, but we've not changed either of these so why do we need to say 'weakens selection' and therefore evolution was affected? --->

Conversely, when pre-copulatory inbreeding allele ($M_{a}$) values were fixed for adaptive inbreeding avoidance, mean $F_{a}$ allele values did not consistently become negative over generations (Figure 4C,D). Fixation of pre-copulatory inbreeding avoidance therefore prevented evolution of post-copulatory inbreeding avoidance. However, in this case, mean $P_{a}$ allele values consistently decreased over generations, reflecting selection against polyandry regardless whether or not post-copulatory inbreeding strategy was costly (Figure 4C,D). Consequently, when $c_{F}=0$, $F_{a}$ allele values had no effect because females were almost exclusively monandrous, resulting in high drift of $F_{a}$ values among replicates (illustrated by wide standard errors, Figure 4C). However, when $c_{F}=0.02$, $F_{a}$ values remained near zero to minimise direct costs. This lack of selection for post-copulatory inbreeding avoidance was driven by a lack of polyandry, and therefore an inability of females to bias fertilisation among multiple mates. When pre-copulatory inbreeding avoidance and polyandry were both fixed ($M_{a}=-10$ and $P_{a}=1$), $F_{a}$ allele values evolved to similarly negative means as $M_{a}$ allele values in Figure 4A,B (see Supporting Information).

@@

<!--- ==================================================================== --->
<!--- BEGIN COMMENTS NOT TO BE TRANSFERRED TO THE WORD DOCUMENT FOR COAUTHORS
<!--- ==================================================================== --->

<!--- Evolution of reproductive strategies involves multiple potentially interacting and degenerate phenotypes. Here we modelled evolution of three reproductive strategies including polyandry, pre-copulatory inbreeding avoidance, and post-copulatory inbreeding avoidance, focusing on three key questions. Using individual-based modelling, we found that evolution of post-copulatory inbreeding avoidance facilitated evolution of polyandry, that evolution of a costly inbreeding avoidance strategy was inhibited when the other inbreeding avoidance strategy was cost free, and that evolution of post-copulatory inbreeding avoidance but not pre-copulatory inbreeding avoidance was precluded when the other inbreeding avoidance strategy was fixed for strong inbreeding avoidance. We thereby conclude that evolution of inbreeding avoidance strategies and associated polyandry are likely to interact in predictable ways, potentially generating new hypotheses concerning evolution of reproductive strategies within and across taxa. --->

<!---  I think these changes have been good ones, and hence that we’re heading along decent lines with the Discussion. However I think it could to with a moderately hefty edit to 1) tidy up some overarching points about what we did near the start, 2) perhaps clarify the structure of the Discussion in relation to the three questions that are now very clearly stated at the end of the Intro, and 3) tighten up the writing a bit (I can tell that the Discussion is a few iterations behind the rest of the manuscript in terms of tight and precise writing, but that’s easily fixed once we get the structure sorted). --->

<!--- Unusually for me (and sorry if this opening was my suggestion from an earlier draft!), I’m wondering whether it’s a good idea to start this broad, or whether we’d be better to keep a bit more focused to start with, and then bring in the generality after that? I’ll make some suggestions, see what you think. Writing is NOT intended to be a definitive version, just me getting some ideas down!  --->

<!--- I do think we need to get this into the start of the Discussion. It’s because our current analyses don’t play around with the magnitude of ID, it’s a fixed parameter = strong. Think we need to make sure that’s clear in peoples’ minds, otherwise we’ll get people asking about IP etc... Hence my suggestion to nail it at the start – which then mirrors the start of the Intro. --->

<!--- Too much/too vague no doubt – but this is a hanger to link back to the bit you flagged up as a worry in the Intro, about whether someone would query why we need ‘dynamic models’ – so here we could flag up the dynamicism again somehow? --->

<!--- To me, this sentence isn’t very helpful, doesn’t say what the questions are or what we concluded, or even how we tackled them, so it seems pretty redundant. Can we weave into next section along lines suggested? --->

<!--- SUGGESTED CHANGE: Strong inbreeding depression in offspring fitness is widely hypothesised to drive evolution of both pre-copulatory and post-copulatory inbreeding avoidance, and hence drive evolution of polyandry [@Zeh1997; @Jennions2000; @Tregenza2002]. However, evolution of such degenerate (i.e. functionally overlapping) mechanisms of inbreeding avoidance cannot be presumed to occur independently, meaning that emerging reproductive strategies are likely to be shaped by relative costs, and might exhibit complex dynamic evolutionary feedbacks across multiple generations. --->

<!--- I was wondering about reiterating a clearer KG here, but actually I think it’s fine without – see comment attached to end of this suggested paragraph. --->

<!--- One of my main comments about the Discussion at the moment is that the really clear structure of the 3 questions that’s now set up at the end of the Intro and followed through the Methods and Results has completely disappeared here. Can we get it back again, a bit? Restating the questions here might be overkill, but yet I’m thinking that a short summary version might be helpful – something like this? --->

<!--- I didn’t get much from this here, and anyway you pretty much repeat the exact same thing in the ‘conclusions’ section at the end? So, if we have to have a conclusions section (for JEB?), then I think it’s best there? --->

<!--- This sentence might need tweaking a bit, but I think something like it would be a good strong end to this first para? --->

<!--- So this is where I start to find the structure of the Discussion a bit hard to follow – seem to be addressing the three questions in reverse order. This first bit (and the link to fig 4), is really addressing question 3, isn’t it? And then we go on to Q2 (fig 3) and then finally Q1 (fig 1). This is pretty confusing. Would it be possible to turn this section around so that the three questions are dealt with in the ‘right’ order? I haven’t tried to implement. Or if you want to keep it like this we might need to clarify that structure and (implicitly or explicitly) note why things are reversed? --->

<!--- ==================================================================== --->
<!--- END COMMENTS NOT TO BE TRANSFERRED TO THE WORD DOCUMENT FOR COAUTHORS
<!--- ==================================================================== --->

Evolution of reproductive strategies cannot be presumed to occur in isolation in wild populations, and is instead likely to involve multiple interacting phenotypes resulting in the potential for feedback and degeneracy (i.e., functional redundance) among phenotypes under selection. For example, strong inbreeding depression in offspring fitness is widely hypothesised to drive selection for inbreeding avoidance, which could consequently feed back to drive selection for polyandry [e.g., @Zeh1997; @Jennions2000; @Tregenza2002]. Evolution of inbreeding avoidance can be driven by selection on both pre-copulatory and post-copulatory mechanisms, so simultaneous evolution of each phenotype is likely to be defined by degeneracy, in addition to any direct costs on phenotypes. We used individual-based modelling to explore evolution of pre-copulatory and post-copulatory inbreeding avoidance and polyandry given (1) hard constraints on evolution of post-copulatory inbreeding avoidance, (2) asymmetric costs of pre-copulatory and post-copulatory inbreeding avoidance, and (3) evolution of one inbreeding avoidance phenotype given the pre-existence the other. We thereby highlight fundamental but theoretically under-developed relationships between evolution of female mating frequency and pre-copulatory versus post-copulatory inbreeding avoidance.

<!--- I have tried to re-work this first paragraph, but I had a very hard time following the logical flow of ideas given the suggested changes. I understand that you want to get all of the components of this manuscript into the front of the Discussion, but I think they need to fit together much more coherently. I didn't quite understand the motivation behind the suggested changes. For example, the proposed first sentence of the Discussion shifted from talking generally about all of our results to only referring to one of our questions. I didn't understand the proposed second sentence at all, which suggested that evolution of degenerate traits can't be assumed to occur independently? (But this is trivially true -- almost by definition! And it inverts the key logic -- the point is that traits don't have independent effects *potentially leading to degeneracy*).  The next part of the proposed second sentence suggested that it necessarily follows from degeneracy that relative costs will shape reproductive interactions, but why does this have to be the case? There needs to be some logical connection between these two claims that just isn't there. I couldn't really figure out how to re-work it because I was really unclear as to what the logical progression of ideas was. Can you clarify what you mean by this? --->

<!--- Is this really correct? Degeneracy is defined by different processes having the same effect, so I don't think pre- and post-copulatory inbreeding avoidance are *affected by* degeneracy so much as *characterised by*, or *an example of* degeneracy. --->

<!--- I think the logic of the first sentence is important here, though it might need some tweaks. The point is that phenotypes don't occur in isolation -> multiple phenotypes interact -> two types of interaction are feedbacks and degeneracy. Hence, degeneracy is one *type* of way that phenotypes might interact, so I think that the sentence structure needs to reflect this so readers don't get confused (i.e., we can't just skip to 'such degenerate' because it doesn't clearly follow from the previous sentence. --->

<!--- I don't think just mentioning 'dynamic models' is enough to make a convincing argument that they were necessary to address the knowledge gap. If we really wanted to flesh out the idea, we would have to explain *why* we couldn't get the answer we did with non-dynamic models. This isn't obvious or trivial! --->

>>>>

Honestly, I also think that some of our wording here is a bit insulting to empirical biologists, many of whom are quite capable of reading, understanding, and applying theory.

@@

However, when both pre-copulatory and post-copulatory inbreeding avoidance could evolve, their relative evolutionary dynamics depended on their relative costs. When pre-copulatory but not post-copulatory inbreeding avoidance was costly ($c_{M}=0.02$ and $c_{F}=0$), pre-copulatory inbreeding avoidance initially evolved (i.e., negative $M_{a}$) but then evolved back towards random mating (i.e., $M_{a}=0$) following increasing evolution of post-copulatory inbreeding avoidance (Figure 3B). Similarly, when post-copulatory but not pre-copulatory inbreeding avoidance was costly ($c_{F}=0.02$ and $c_{M}=0$), post-copulatory inbreeding avoidance initially evolved ($F_{a}=0$) before evolving back to random fertilisation after ca $20000$ generations (Figure 3D).

When pre-copulatory inbreeding avoidance was costly, allowing evolution of cost free post-copulatory inbreeding avoidance greatly facilitated evolution of polyandry (Figure 3A versus 3B). However when post-copulatory inbreeding avoidance was costly, allowing evolution of cost free pre-copulatory inbreeding avoidance caused $P_{a}$ alleles to decrease to very negative values, reducing expression of polyandry (Figure 3C versus 3D; polyandry was cost free in all these simulations).

@@

Conversely, when pre-copulatory inbreeding allele ($M_{a}$) values were fixed for adaptive inbreeding avoidance, mean $F_{a}$ allele values did not consistently become negative over generations (Figure 4C,D). Fixation of pre-copulatory inbreeding avoidance therefore prevented evolution of post-copulatory inbreeding avoidance. However, in this case, mean $P_{a}$ allele values consistently decreased over generations, reflecting selection against polyandry regardless whether or not post-copulatory inbreeding avoidance was costly (Figure 4C,D). Consequently, when $c_{F}=0$, $F_{a}$ allele values had no effect because females were almost exclusively monandrous, resulting in high drift of $F_{a}$ values among replicates (illustrated by wide standard errors, Figure 4C). However, when $c_{F}=0.02$, $F_{a}$ values remained near zero to minimise direct costs. This lack of selection for post-copulatory inbreeding avoidance was driven by a lack of polyandry, and therefore an inability of females to bias fertilisation among multiple mates. When pre-copulatory inbreeding avoidance and polyandry were both fixed ($M_{a}=-10$ and $P_{a}=1$), $F_{a}$ allele values evolved to similarly negative means as $M_{a}$ allele values in Figure 4A,B (see Supporting Information).

@@

Evolution of reproductive strategies cannot be presumed to occur in isolation from each other in real-world populations, and is instead likely to involve multiple interacting phenotypes resulting in the potential for feedback and degeneracy (i.e., functional redundance) among phenotypes under selection. For example, strong inbreeding depression in offspring fitness is widely hypothesised to drive selection for inbreeding avoidance, which could consequently feed back to drive selection for polyandry [e.g., @Zeh1997; @Jennions2000; @Tregenza2002]. Evolution of inbreeding avoidance can be driven by selection on both pre-copulatory and post-copulatory mechanisms, so simultaneous evolution of each phenotype is likely to be defined by degeneracy, in addition to any direct costs on phenotypes. We used individual-based modelling to explore evolution of pre-copulatory and post-copulatory inbreeding avoidance and polyandry given (1) hard constraints on evolution of post-copulatory inbreeding avoidance, (2) asymmetric costs of pre-copulatory and post-copulatory inbreeding avoidance, and (3) evolution of one inbreeding avoidance phenotype given the pre-existence the other. We thereby highlight fundamental but theoretically under-developed relationships between evolution of female mating frequency and pre-copulatory versus post-copulatory inbreeding avoidance.
